---
title: 'Lab Assignment #3 - Recap'
author: "Heike Hofmann"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, middle, center
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, echo = TRUE)
```
# A systems engineer walks into a bar ...

---
class: inverse
# ... 

- he orders a beer

- he orders 9,999,999 beers

- he orders -1 beers

- he orders C5F63A beers

- he orders a lizard

---
class: inverse
# the first real customer enters the bar

... and asks where the bathroom is.

--

<br> <br> <br>

The bar catches fires, explodes and kills everyone 💥💥💥

---

# Submission

- something went wrong with Canvas submissions - groups were permuted??

- please submit your team submission again under `Lab 3 2nd attempt`


---

# What we will do to your package

1. download repo from github 

2. install and run checks 

3. run your defined tests: `devtools::test()`

4. run an example:

```{r eval = FALSE}
library(tidyverse)
library(pkg_name)
oz <- team_X("gadm36_AUS_shp/gadm36_AUS_1.shp")
oz %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = NAME_1))
```

5. run your examples (I will come back to that)

6. visit the package's website at https://team_member.github.io/pkg_name/

6. run some of our checks 

---

# Why do we do this to you?

- expectation: all of the previous checks work without problems.
As always, when you encounter a problem, make a note of it and put it into the Readme file. 

- this lab is a technical test run for the final project

- new due date: Suntoday,r 31 25 p1:59


---

# Final project

submit topic for the final project by Wednesday Mar 27.


---

# Your examples

Where to place the data for your example?

For your example you will need a set of five files: cpg, dbf, prj, shp, and shx

Non-R data are being placed in the folder `inst`

During the installation all elements of `inst` are placed in a special location on your machine.

For example, to find `inst/extdata/mydata.csv`, you'd call `system.file("extdata", "mydata.csv", package = "mypackage")`

---

# Example of an example

Assume you have the australia files in a folder called `extdata` in folder `inst`

The example then should be:
```{r eval=FALSE}
ozfilepath <- system.file("extdata/gadm36_AUS_1.shp", ".shp", package = "mypackage")
ozdata <- team_X(ozfilepath, tolerance=0.1)
```
