---
title: 'Lab Assignment #3: Wrapping the world in a package'
author: "Heike Hofmann"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, echo = TRUE)
```
# Outline

- Creating an R package

- Cleaning up last lab's code and wrapping it in functions

- Writing some tests and documentation


<br><br><br>

The final deliverable is again a self-contained RMarkdown file.

We will install your package, read through the documentation, run some examples and run your tests on it (we might have some tests of our own as well).


---

# Lab organization

1. In RStudio, create a new package as a team, i.e. only one of you should create the package. Make sure that the package is not yet used as an R package on CRAN. 

2. Create a corresponding github repository.  Make sure to create a README.Rmd file, that keeps the md file (see yaml options).
Make all of the team members contributors (under `Settings` > `Collaborators`)

3. Include a link to the repository in the README.Rmd file


---

# Last time ...

... you were asked to write a function that allowed you to create a data frame from 
a `geometry` variable

# This time:

- each team member is asked to pick the solution to last time's lab (listed in the back) from a different team (and not the team they were on)

- wrap the solution into a function of the form

`team_X <- function(file, tolerance)` where `X` is the solution for team X used,  file is a file path to a shape file and 
tolerance is the value used for thinning the polygon


The return value is a data frame of the geographic information of the polygons and the additional information (such as name of the country, name of the territory/state, ...)

---

# Documentation

Using roxygen tags, appropriately document function `team_X`.

# Checks

Include checks on elements of your function to ensure your function cannot be broken easily and gives appropriate warning or error messages.

# Testing

Write tests to ensure your function is working as intended.

---

# Deliverable

**Upload your README.Rmd file to Canvas. Only one submission per team is required. 
Finishing touches can be made until Sunday, Mar 24, 10 pm.**

---

# Necessary preparation for any solution

```{r dependencies, warning = FALSE, message = FALSE}
library(tidyverse)
library(purrr)
library(sf)

ozbig <- read_sf("./data/gadm36_AUS_shp/gadm36_AUS_1.shp")
oz_st <- maptools::thinnedSpatialPoly(as(ozbig, "Spatial"), tolerance = 0.1, minarea = 0.001, topologyPreserve = TRUE)
oz <- st_as_sf(oz_st)

```

---

## Team 1

```{r}
df.oz.purr <- oz$geometry %>%
  map_depth(3, data.frame) %>%
  purrr::flatten() %>%
  purrr::flatten() %>%
  bind_rows(.id = "group") %>%
  rename("lat" = y, "long" = x)

# df.oz.purr %>% 
#   ggplot(aes(x = long, y = lat, group = group)) + 
#     geom_polygon()
```

---

## Team 2

```{r}
## Help calculate the group.
## Input: oz$geometry
helper.group <- function(geo){
  geo %>% flatten() %>% flatten()-> dd
  countgrouprep <- flatten_int(map(dd, nrow))
  num_group <- length(countgrouprep)
  rep(1:num_group, time = countgrouprep)
}

## Help calculate the order.
## Input: oz$geometry[[i]]
helper.order <- function(geol){
  geol %>% flatten() -> d
  longlat <- do.call(rbind, d)
  order_num <- sum(flatten_int(map(d, nrow)))
  order <- seq(1:order_num)
  cbind(longlat, order)
}
```

---

```{r}
## Converts shapefile to lat-long file
## Input: oz$geometry
sh2lat <- function(geofile){
  res <- map(geofile, .f=helper.order)
  ress <- do.call(rbind, res)
  group <- helper.group(geofile)
  ress <- cbind(ress, group)
  colnames(ress) <- c("long", "lat", "order", "group")
  ress <- as.data.frame(ress)
}

ozplus <- sh2lat(oz$geometry)
#ozplus %>% ggplot(aes(x=long,y=lat,group=group)) + 
#  geom_polygon() + theme_bw()
```

---

## Team 3

```{r}
drawmap <- function(dat) {
  con_big <- read_sf(dat)
  con_st <- maptools::thinnedSpatialPoly(
    as(con_big, "Spatial"), tolerance = 0.1, 
    minarea = 0.001, topologyPreserve = TRUE)
  con <- st_as_sf(con_st)
  con_plus <- con$geometry %>% 
    flatten() %>% 
    flatten() %>% 
    lapply(., as.data.frame) %>% 
    bind_rows() %>% 
    rename(., c("x" ="long", "y" = "lat"))
  temp <- con$geometry %>% 
    flatten() %>% 
    flatten() %>% 
    lapply(., nrow) %>% 
    unlist
  con_plus$group <- rep(1 : length(temp), temp)

  con_plus %>% 
    ggplot(aes(x = long, y = lat, group = group)) + 
    geom_polygon()
}
```

---

```{r, eval = FALSE}
# drawmap(dat = "data/gadm36_AUS_shp/gadm36_AUS_1.shp")
```

---

## Team 4

```{r}
poly2df <- function(feature, unlist = T) {
  if (unlist) feature <- unlist(feature, recursive = F)
  lapply(feature, function(x) {
    mutate(rename_all(data.frame(x), ~ c("long", "lat")), order = row_number())
  }) %>% tibble(polygon = .)
}

# create new dataset reforming geometry to long, lat, order, group
makedata <- function(data, thin = T, unlist = T) {
  if (thin) {
    oz_st <- maptools::thinnedSpatialPoly(as(data, "Spatial"), tol = 0.1, min = 0.001, topologyPreserve = T)
    oz <- st_as_sf(oz_st)  # install package: rgeos
  } else oz <- data
  
  as_tibble(oz) %>% mutate(new = map(geometry, poly2df, unlist = unlist)) %>% select(-geometry) %>%
    unnest() %>% mutate(group = row_number()) %>% unnest()
}
```

---

```{r}
australia <- read_sf("data/gadm36_AUS_shp/gadm36_AUS_1.shp")
australia_plus <- makedata(australia)
# australia_plus %>%
#   ggplot(aes(x = long, y = lat, group = group)) +
#   geom_polygon(color = "black", fill = "white", size = 0.2) +
#   labs(x = "Longitude", y = "Latitude", title = "Australia") +
#   coord_fixed() +
#   theme_bw() +
#   theme(plot.title = element_text(hjust = 0.5))
```

---

## Team 5

```{r}
Mat2Df <- function(Mat){
  long <- Mat[,1]
  lat <- Mat[,2]
  order <- 1:nrow(Mat)
  group <- rep(rnorm(1),nrow(Mat))
  df <- data.frame(long=long,lat=lat,group=group,order=order)
  df
}
oz_flatten <- flatten(flatten(oz$geometry))
ozplus <- purrr::map_df(.x=oz_flatten,.f=Mat2Df)
#ozplus %>% ggplot(aes(x=long, y=lat, group=group)) + 
#  geom_polygon()
```

---

## Team 6

```{r}
mat2df <- function(mat) {
  df <- data.frame(long = mat[, 1],
                   lat  = mat[, 2], 
                   temporary.group = sum(mat[,1]+mat[,2] *nrow(mat) + rnorm(1, 100, 10)), 
                   order= 1:nrow(mat))
  df
}
```

```{r}

# flatten oz$geometry twice to obtain a list of matrices
# this is because we have two layers of lists ahead of matrices
oz.geometry.flattened <- oz$geometry %>% flatten() %>% flatten()

# Create a dataframe ozplus from the geometry variable using purrr
# map_df applies a function to elements of a list and bind the dataframes together
# Instead of temporary.group variable by the function mat2df, 
# group variable consisting of the index corresponding each matrix will be used.
ozplus <- oz.geometry.flattened %>% purrr::map_df(.x = ., .id ="group", .f = mat2df)

# plot the result
# ozplus %>% ggplot(aes(x = long, y = lat, group = group)) + geom_polygon()
```

---

## Team 7

```{r}
f <- function(dframe){
  dframe <- data.frame(order = c(1:nrow(dframe)), long = dframe$x, lat = dframe$y)
}

## Here our depth is 3, but that could change depending on the file
ozplus <- oz$geometry %>% 
  modify_depth(3,data.frame) %>% 
  modify_depth(3,f) %>%
  flatten() %>%
  flatten() %>%
  bind_rows(.id = "group")

# ozplus %>% 
#   ggplot(aes(x = long, y = lat, group = group)) + 
#   geom_polygon(fill = "white", colour = "black", lwd = 1)+
#   theme_bw()+
#   coord_quickmap()
```

---

## Team 8

```{r}
# Helper fucntion to turn one of these matrices into a single data frame with variables long, lat, group, order. 
helper <- function(y) {
  data <- data.frame(long = y[, 1],
                     lat  = y[, 2], 
                     groups = rep(rnorm(1),nrow(y)),
                     order = 1:nrow(y))
  data
}

geometry <- oz$geometry

# In order to receive list of matrices, we used the flatten 
matrix <- geometry %>% flatten() %>% flatten()

# By using  purrr functionality, we create the dataframe
ozplus <- matrix %>% purrr::map_df(.x = ., .id ="group", .f = helper)

# Plot the map
#ozplus %>% ggplot(aes(x = long, y = lat, group = group)) + geom_polygon()
```

---

## Team 9

this slide does not show the whole solution - make sure to go to the Rmd file on github to get the complete solution
```{r, echo = -21:-35}
reoganise <- function(x) {
  ### i, j, k represent the index of the geometry[[i]][[j]][[k]])
  ### work on k
  reorganise_k <- function(x) {
    force(x)
    y <- data.frame(x, group = group, order = 1:nrow(x), i = i, j = j, k = k)
    group <<- group + 1
    k <<- k + 1
    return(y)
  }
  
  ### work on j
  reorganise_j <- function(x) {
    force(x)
    k <<- 1
    y <- lapply(x, FUN = reorganise_k)
    j <<- j + 1
    return(y)
  }
  
  ### work on i
  reorganise_i <- function(x) {
    force(x)
    j <<- 1
    y <- lapply(x, FUN = reorganise_j)
    i <<- i + 1
    return(y)
  }
  
  # set initial value
  group <- 1
  i <- 1
  j <- 1
  k <- 1
  
  # run
  geometry_my1 <- lapply(x, reorganise_i)
  
  reducetodata <- function(x) {
    lapply(x, Reduce, f = rbind)
  }
  
  temp1 <- lapply(geometry_my1, reducetodata)
  temp2 <- lapply(temp1, Reduce, f = rbind)
  ozplus <- Reduce(rbind, temp2)

  names(ozplus) <- c("long", "lat", "group", "order", "i", "j", "k")
  return(ozplus)
}
```

---

```{r}
oz_reorg <- reoganise(oz$geometry)

# oz_reorg %>% ggplot(aes(x = long, y = lat, group = group)) + 
#   geom_polygon()
```

---

## Team 10

```{r}
shpBigToSmall <- function(dsn) {
  
  shpbig <- read_sf(dsn)
  shp_st <- maptools::thinnedSpatialPoly(
    as(shpbig, "Spatial"), tolerance = 0.1, 
    minarea = 0.001, topologyPreserve = TRUE)
  shp <- st_as_sf(shp_st)
  
  shpSmall <- shp %>% select(NAME_1, geometry) %>% 
    group_by() %>% 
    mutate(coord = geometry %>% map(.f = function(m) flatten(.x=m)),
           region = row_number()) %>% 
    unnest
  st_geometry(shpSmall) <- NULL
  shpSmall <- shpSmall %>% 
    mutate(coord = coord %>% map(.f = function(m) as_tibble(m)),
                                group = row_number()) %>% 
    unnest %>% 
    setNames(c("name", "region","group", "long", "lat"))

  return(shpSmall) 
}
```

---

```{r}
dsn <- "data/gadm36_AUS_shp/gadm36_AUS_1.shp"
shpSmall <- shpBigToSmall(dsn)
plotMap <- shpSmall %>% ggplot(aes(x = long, y = lat, group = group)) + 
  geom_polygon(color = "black")
#print(plotMap)
```

---

## Team 11

```{r}
# This function extracts data from a shapefile with 2 levels in spf$geometry
extract_data_L2 <- function(spf) {
# Reads in the shapefile and extracts data from nested lists, finally
# recording them into one data frame with columns: long, lat, group and
# order.
#
# Args:
#   spf:    shapefile of a country
#
# Output:
#   df:     data frame
    
    map_depth(.x = spf$geometry, 2, .f = c(1)) %>% flatten %>% 
        map_dfr(data.frame, .id = "group") -> df 
    
    # rename columns
    colnames(df) <- c("group", "long", "lat")
    
    # add new variable = "order"
    df$order <- seq(from = 1, to = nrow(df), by = 1)
    
    return(df)
}
```

---

```{r}
ozplus <- extract_data_L2(oz)

# # plot
# ozplus %>% ggplot(aes(x = long, y = lat, group = group, na.rm = TRUE)) + geom_polygon(colour = "black", fill = "grey50")
```

---

## Team 12

```{r}
helper <- function(d){
  d <- unlist(d,recursive = FALSE)
  d <- purrr::map(d,.f=add_order)
  d <- add_layer(d)
  return(d)
}


add_order <- function(d){
  l <- nrow(d)
  return(cbind(d,seq(1,l,by=1)))
}
add_layer <- function(d){
  ll <- unlist(lapply(d,nrow))
  d <- do.call(rbind,d)
  d <- cbind(d,rep(c(1:length(ll)),time=ll))
  return(d)
}
```
---

```{r}
purrr::map(oz$geometry, .f=helper) -> res

res <- add_layer(res)
colnames(res) <- c('long','lat','order','group','geo')
res <- as.data.frame(res)

# ggplot(data=res)+
#   geom_path(aes(x=long,y=lat,
#                 group=paste(res$geo,res$group,sep='.'))) + 
#   theme_bw()
```
